Modeling Two-Dimensional Vesicle Hydrodynamics with Hydrophobic Attraction Potential Simulations

### 2021-08-06 doc/main_jfm.tex
finished up first run at secs 2 and 3

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\uu}{\mathbf{u}}
\renewcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}

@@

%%%
\section{Integral Equation Method}
\label{sec:IEM}
Since the fluid velocity is assumed to be Stokesian, it can be written
in terms of layer potentials. Following previous work of~\cite{Lukas19},
we represent the velocity as the sum of a double-layer potential and the
$N_b$ Stokeslets and Rotlets
\begin{align}
  \label{eqn:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
The double-layer potential is
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nnu}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy,
\end{align}
where $\rr = \xx - \yy$ and $\rho = |\rr|$. The Stokeslet and Rotlet
centered at $\aa_i$ are
\begin{align}
  \SS(\xx,\aa_i)\FF_i &= \frac{1}{4\pi} \left(-\log(\rho) + 
    \frac{\rr \otimes \rr}{\rho^2} \FF_i \right), \\
  \RR(\xx,\aa_i)T_i &= \frac{1}{4\pi} \frac{\rr^\perp}{\rho^2} T_i,
\end{align}
where $\rr = \xx - \aa_i$ and $\rho = |\rr|$. The Stokeslet is
torque-free and has force $\FF_i$ while the Rotlet force-free and has
torque $T_i$. Therefore, the velocity~\eqref{eqn:velocity} satisfies the
total force~\eqref{eqn:force} and torque~\eqref{eqn:torque} conditions
if the double-layer potential $\DD[\eeta]$ is force- and torque-free.
Matching the limit of~\eqref{eqn:velocity} with the rigid body motion,
and imposing that $\DD[\eeta]$ is force- and torque-free, the density
function $\eeta$, translational velocity $\vv_i$, and angular velocity
$\omega_i$ satisfy
\begin{alignat}{3}
  \nonumber
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eqn:SKIE}
    + \sum_{j=1}^{N_b} &
    \left(\SS(\xx,\aa_j) \FF_j + \RR(\xx,\aa_j) T_j\right),
    \quad &&\xx \in \bd \Gamma_i,\: i=1,\ldots,N_b, \\
  \label{eqn:mobility1}
  \int_{\bd\Gamma_i} \eeta \cdot \nnu_i \, ds &= {\bf 0}, 
  &&i = 1,\ldots,N_b, \\
  \label{eqn:mobility2}
  \int_{\bd\Gamma_i} \eeta\times(\xx-\aa_i)^\perp \cdot \nnu ds &= 0,
  &&i = 1,\ldots,N_b.
\end{alignat}

%Given a collection of particles with forces and torques calculated from
%the hydrophobic interactions and the short range repulsion,
%equations~\eqref{eqn:SKIE}--\eqref{eqn:mobility2} form a solvable set of
%integral equations for the translational velocities $\vv_i$, the angular
%velocities $\omega_i$, and the density function $\eeta$.

We discretize each of the $\Gamma_i$, $i=1,\ldots,N_b$, at $N$
discretization points. The trapezoid rule is used to approximate each of
the integrals in equations~\eqref{eqn:mobility1}
and~\eqref{eqn:mobility2}, and the double-layer potential
equation~\eqref{eqn:SKIE}. When two bodies are close, the double-layer
potential becomes nearly-singular, and we replace the double-layer
potential with an interpolation-based nearly-singular quadrature
rule~\cite{qua-bir2014}. After discretizing and applying appropriate
quadrature rules, the result is a $(2NN_b + 3N_b) \times (2NN_b + 3N_b)$
linear system whose condition number is independent of $N$, and this
linear system is solved using matrix-free GMRES. To reduce the number of
GMRES iterations, a block-diagonal preconditioner is used.

@@

\section{Appendices}
The stress ${\bf T}$ contains the gradient of the double layer potential. This 
introduces an inherent difficulty in evaluating stress on the boundary.
To overcome the difficulty of evaluation on $\Sigma_p$, for instance, we decompose $u$ into singular and nonsingular parts:
\begin{equation}
u = u_p + v_p
\end{equation}
where  
\begin{equation}
u_p(x) = \int_{\Sigma_p} \frac{\partial G(x,y)}{\partial \nnu} \eta(y) \,\dif S(y).
\end{equation}
It turns out that the contribution to the force and torque coming from the singular
part vanishes. In particular, we will prove the identity
\begin{equation}
\label{eq:recipforcetorque}
F_p + iG_p= \int_{\Sigma_p} \gamma ({\bf I}+i{\bf X})J_{p} \,\dif S
\end{equation}
where
 \begin{equation}
\label{eq:jumpstress1}
J_{p} = 2\rho^{-1} \eta  v_p \nnu 
+ 2\rho \nabla \eta \cdot \tau_i(\nabla v_p \cdot \tau_i \nnu -  \nabla
   v_p \cdot \nnu \tau_i).
\end{equation}
Here $\nnu$ is the unit normal, $\{\tau_i\}_{i=1}^2$ are orthonormal tangent vectors,
and ${\bf X}$ is the skew symmetric matrix with axial vector $x$, i.e.
\[X_{ij} = \varepsilon_{ikj}x_k,\quad {\bf X}v = x\times v\]
where $\boldsymbol{\varepsilon}$ is the third-order alternating tensor.
We employ the Einstein summation convention and use complex vectors to streamline the presentation, $i^2 = -1$.

>>>>

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\uu}{\mathbf{u}}
\renewcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}

@@

%%%
\section{Integral Equation Method}
\label{sec:IEM}
Computing the hydrophobic attraction potential and the particle forces
and torques requires solving an elliptic partial differential equation
(PDE) in an unbounded complex domain. We recast both these PDEs as
boundary integral equations (BIEs). We discretize each BIE at $N$ points
on each of the $N_b$ particles with a collocation method. Integrals that
are smooth are computed with the spectrally-accurate quadrature rule,
and nearly-singular integrals, caused by close contact between two
particles, are computed with a high-order interpolation-based quadrature
rule~\cite{qua-bir2014}. After discretizing and applying quadrature, the
resulting linear system is solved with matrix-free GMRES, and we
guarantee that the number of GMRES iterations is mesh-independent by
using second-kind BIEs.

\subsection{HAP Boundary Integral Equation}
Similar to our previous work~\cite{Fu20}, we represent the HAP as a
double-layer potential
\begin{align}
  \label{eqn:HAP_DLP}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, ds_\yy,
    \quad \xx \in \Omega,
\end{align}
where $K_0$ is the zeroth-order modified Bessel function of the first
kind. By requiring that the density function $\sigma$ satisfies the
second-kind integral equation
\begin{align}
  f(\xx) = \frac{1}{2}\sigma(\xx) + 
    \frac{1}{2\pi}\int_{\bd\Omega} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, ds_\yy,
    \quad \xx \in \bd\Omega,
\end{align}
the HAP double-layer potential~\eqref{eqn:HAP_DLP} satisfies the
screened Laplace equation~\eqref{SL}--\eqref{SLff}. To calculate the
hydrophobic force and torque, the gradient of the double-layer
potential~\eqref{eqn:HAP_DLP} must be computed on the boundary of each
particle. In Appendix~\ref{sec:appendixA}, we show that the force and
torque calculations can be expressed in terms of integral that are
non-singular.

\subsection{Mobility Problem Boundary Integral Equation}
Following previous work of~\cite{Lukas19}, we represent the velocity as
the sum of a double-layer potential and the $N_b$ Stokeslets and Rotlets
\begin{align}
  \label{eqn:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
The double-layer potential is
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nnu}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy,
\end{align}
where $\rr = \xx - \yy$ and $\rho = |\rr|$. The Stokeslet and Rotlet
centered at $\aa_i$ are
\begin{align}
  \SS(\xx,\aa_i)\FF_i &= \frac{1}{4\pi} \left(-\log(\rho) + 
    \frac{\rr \otimes \rr}{\rho^2} \FF_i \right), \\
  \RR(\xx,\aa_i)T_i &= \frac{1}{4\pi} \frac{\rr^\perp}{\rho^2} T_i,
\end{align}
where $\rr = \xx - \aa_i$ and $\rho = |\rr|$. The Stokeslet is
torque-free and has force $\FF_i$ while the Rotlet force-free and has
torque $T_i$. Therefore, the velocity~\eqref{eqn:velocity} satisfies the
total force~\eqref{eqn:force} and torque~\eqref{eqn:torque} conditions
if the double-layer potential $\DD[\eeta]$ is force- and torque-free.
Matching the limit of~\eqref{eqn:velocity} with the rigid body motion,
and imposing that $\DD[\eeta]$ is force- and torque-free, the density
function $\eeta$, translational velocity $\vv_i$, and angular velocity
$\omega_i$ satisfy
\begin{alignat}{3}
  \nonumber
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eqn:SKIE}
    + \sum_{j=1}^{N_b} &
    \left(\SS(\xx,\aa_j) \FF_j + \RR(\xx,\aa_j) T_j\right),
    \quad &&\xx \in \bd \Gamma_i,\: i=1,\ldots,N_b, \\
  \label{eqn:mobility1}
  \int_{\bd\Gamma_i} \eeta \cdot \nnu_i \, ds &= {\bf 0}, 
  &&i = 1,\ldots,N_b, \\
  \label{eqn:mobility2}
  \int_{\bd\Gamma_i} \eeta\times(\xx-\aa_i)^\perp \cdot \nnu ds &= 0,
  &&i = 1,\ldots,N_b.
\end{alignat}

The trapezoid rule is used to approximate each of the integrals in
equations~\eqref{eqn:mobility1} and~\eqref{eqn:mobility2}, and the
double-layer potential equation~\eqref{eqn:SKIE} is approximated with
either the trapezoid rule, or the interpolation-based near-singular
integration quadrature rule. After discretizing and applying appropriate
quadrature rules, the result is a $(2NN_b + 3N_b) \times (2NN_b + 3N_b)$
linear system that we precondition with a block-diagonal preconditioner.
We note that other BIE formulations of the mobility problem use
single-layer potentials~\cite{cor-gre-rac-vee2017, rac-gre2016} or both
single- and double-layer potentials~\cite{cor-vee2018}.

@@

\section{Appendices}
\label{sec:appendixA}
The stress ${\bf T}$ contains the gradient of the double layer
potential. This introduces an inherent difficulty in evaluating stress
on the boundary.  To overcome the difficulty of evaluation on
$\Sigma_p$, for instance, we decompose $u$ into singular and nonsingular
parts:
\begin{equation}
u = u_p + v_p
\end{equation}
where  
\begin{equation}
u_p(x) = \int_{\Sigma_p} \frac{\partial G(x,y)}{\partial \nnu} \eta(y) \,\dif S(y).
\end{equation}
It turns out that the contribution to the force and torque coming from the singular
part vanishes. In particular, we will prove the identity
\begin{equation}
\label{eq:recipforcetorque}
F_p + iG_p= \int_{\Sigma_p} \gamma ({\bf I}+i{\bf X})J_{p} \,\dif S
\end{equation}
where
 \begin{equation}
\label{eq:jumpstress1}
J_{p} = 2\rho^{-1} \eta  v_p \nnu 
+ 2\rho \nabla \eta \cdot \tau_i(\nabla v_p \cdot \tau_i \nnu -  \nabla
   v_p \cdot \nnu \tau_i).
\end{equation}
Here $\nnu$ is the unit normal, $\{\tau_i\}_{i=1}^2$ are orthonormal tangent vectors,
and ${\bf X}$ is the skew symmetric matrix with axial vector $x$, i.e.
\[X_{ij} = \varepsilon_{ikj}x_k,\quad {\bf X}v = x\times v\]
where $\boldsymbol{\varepsilon}$ is the third-order alternating tensor.
We employ the Einstein summation convention and use complex vectors to streamline the presentation, $i^2 = -1$.
